An Improved Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium

### 2021-07-13 simplified_sdc_intro/paper.tex
some more Jac work

The reaction vector is
\begin{equation}
\Rb(\Uc^\prime) = \left (  \begin{array}{c} 0 \\ \rho \omegadot_\alpha \\ \rho \omegadot_\beta \\ \rho \Sdot \\ \rho \Sdot \end{array} \right )
\end{equation}
and the Jacobian is computed as $\partial \Rb/\partial {\bf w}$:
\begin{equation}
\frac{\partial \Rb}{\partial {\bf w}} = \left (
  \begin{array}{ccccc}
     0 & 0 & 0 & 0 & 0 \\
     \omegadot_\alpha + \rho \frac{\partial \omegadot_\alpha}{\partial \rho} &
     \rho \frac{\partial \omegadot_\alpha}{\partial X_\alpha} &
     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & 0 & \rho \frac{\partial \omegadot_\alpha}{\partial T} \\
      %
     \omegadot_\beta + \rho \frac{\partial \omegadot_\beta}{\partial \rho} &
     \rho \frac{\partial \omegadot_\beta}{\partial X_\alpha} &
     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} & 0 &  \rho \frac{\partial \omegadot_\beta}{\partial T} \\
     %
     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial T} \\
     %
     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial T} \\
  \end{array}
  \right )
\end{equation}

While VODE can compute the entire Jacobian, ${\bf J}$ numerically via differencing, we found
that does not give reliable results.  Instead, we compute $\partial {\bf w}/\partial {\bf U}^\prime$ analytically and $\partial {\bf R}/\partial {\bf w}$ numerically via one-sided differencing, following
the algorithm in \cite{lsode} to minimize numerical noise.
The resulting Jacobian, as computed via equation~\ref{eq:factored_jac}, is larger than
expected by the integrator because of the inclusion of the $\rho$
terms in the intermediate steps, so we remove the row corresponding to
$R(\rho)$ and the column corresponding to $\partial/\partial \rho$
before passing the Jacobian to VODE.

>>>>

The reaction vector is
\begin{equation}
\Rb(\Uc^\prime) = \left (  \begin{array}{c} 0 \\ \rho \omegadot_\alpha \\ \rho \omegadot_\beta \\ \rho \Sdot \\ \rho \Sdot \end{array} \right )
\end{equation}
and the Jacobian is computed as $\partial \Rb/\partial {\bf w}$:
\begin{equation}
\frac{\partial \Rb}{\partial {\bf w}} = \left (
  \begin{array}{ccccc}
     0 & 0 & 0 & 0 & 0 \\
     \omegadot_\alpha + \rho \frac{\partial \omegadot_\alpha}{\partial \rho} &
     \rho \frac{\partial \omegadot_\alpha}{\partial X_\alpha} &
     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & 0 & \rho \frac{\partial \omegadot_\alpha}{\partial e} \\
      %
     \omegadot_\beta + \rho \frac{\partial \omegadot_\beta}{\partial \rho} &
     \rho \frac{\partial \omegadot_\beta}{\partial X_\alpha} &
     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} & 0 &  \rho \frac{\partial \omegadot_\beta}{\partial e} \\
     %
     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial e} \\
     %
     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial e} \\
  \end{array}
  \right )
\end{equation}

While VODE can compute the entire Jacobian, ${\bf J}$ numerically via differencing, we found
that does not give reliable results.  Instead, we compute $\partial {\bf w}/\partial {\bf \Uc}^\prime$ analytically and $\partial {\bf R}/\partial {\bf w}$ numerically via one-sided differencing, following
the algorithm in \cite{lsode} to minimize numerical noise.

The final Jacobian is found by multiplying these two and removing the row corresponding to
$R(\rho)$ and the column corresponding to $\partial/\partial \rho$:
\begin{equation}
\frac{\partial \Rb}{\partial {\bf \Uc}} = \left (
  \begin{array}{ccccc}
    \frac{\partial \omegadot_\alpha}{\partial X_\alpha} -\frac{\partial \omegadot_\alpha}{\partial e} e_{X_\alpha} &
    \frac{\partial \omegadot_\alpha}{\partial X_\beta} -\frac{\partial \omegadot_\alpha}{\partial e}  e_{X_\beta} &
    0 &
    \frac{\partial \omegadot_\alpha}{\partial e} \\
     %
    \frac{\partial \omegadot_\beta}{\partial X_\alpha} -\frac{\partial \omegadot_\beta}{\partial e} e_{X_\alpha} &
    \frac{\partial \omegadot_\beta}{\partial X_\beta} -\frac{\partial \omegadot_\beta}{\partial e}  e_{X_\beta} &
    0 &
    \frac{\partial \omegadot_\beta}{\partial e} \\
     %
     \frac{\partial \Sdot}{\partial X_\alpha} - \frac{\partial \Sdot}{\partial e} e_{X_\alpha} &
     \frac{\partial \Sdot}{\partial X_\beta} - \frac{\partial \Sdot}{\partial e} e_{X_\beta} &
     0 &
     \frac{\partial \Sdot}{\partial e} \\
     %
     \frac{\partial \Sdot}{\partial X_\alpha} - \frac{\partial \Sdot}{\partial e} e_{X_\alpha} &
     \frac{\partial \Sdot}{\partial X_\beta} - \frac{\partial \Sdot}{\partial e} e_{X_\beta} &
     0 &
     \frac{\partial \Sdot}{\partial e} \\
  \end{array}
  \right )
\end{equation}
